---
title: "Data Cleaning"
author: "YuetongLiu"
date: '2020-02-08'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install, include=FALSE}
# install.packages(c("dplyr","here","readxl"))
library(dplyr)
library(here)
library(readxl)
library(here)
```

## Load Data

```{r load}
assessment.1620 <- read.csv(here("data/Assessment", "2016 - 2020 Raw.csv"),header = F)
col.name <- read_excel(here("data/Assessment", "2020 Only.xlsx"))
colnames(assessment.1620) <- colnames(col.name)
```

## Filter Data
```{r filter}
municipality.list = c(
  "Burnaby", 
  "Coquitlam", 
  "Delta", 
  "Langley - City", 
  "Langley - Township",
  "Maple Ridge",
  "Maple Ridge Rural", 
  "North Vancouver - City",
  "North Vancouver - Dist",
  "Pitt Meadows", 
  "Port Coquitlam", 
  "Port Moody", 
  "Richmond", 
  "Surrey", 
  "Vancouver", 
  "White Rock", 
  "West Vancouver", 
  "Bowen Island", 
  "Anmore", 
  "Belcarra",
  "Lions Bay",
  "New Westminster")

# Select only relevent features
assessment.thin <- distinct(assessment.1620) %>%
  select(PIC,Year,AssessorAreaCode:AddressAssessorMunicipalityDesc, TaxClassCode:TaxClassTaxRate)%>% 
  filter(TaxClassCode %in% c("01","05","06"),
         AddressAssessorMunicipalityDesc %in% municipality.list,
         Year %in% c(2016,2017,2018,2019,2020))

# Overview
head(assessment.thin)
dim(assessment.thin)
str(assessment.thin)

assessment.thin[assessment.thin == "NULL"] <- NA
assessment.thin$TaxClassTaxRate <- as.numeric(as.character(assessment.thin$TaxClassTaxRate))
```

## Missing Value
```{r missing}
# Check Missing Value 
for (feature in colnames(assessment.thin)){
  variable = eval(parse(text = paste0("assessment.thin$",feature)))
  null.count = sum(is.na(variable))
  print(paste(feature,null.count, sep = ":"))
} 
```

## Check if TaxClassTaxRate is constant at same Year, AddressAssessorMunicipalityDesc, TaxClassCode
```{r}
(TaxClassTaxRate.text <- assessment.thin%>%
   group_by(Year, AddressAssessorMunicipalityDesc,TaxClassCode)%>%
   filter(!is.na(TaxClassTaxRate))%>%
   summarise(test = var(as.numeric(as.character(TaxClassTaxRate))))%>%
   filter(test !=0))


Langley <- assessment.thin%>%
  filter(Year %in% c(2019, 2020),
         AddressAssessorMunicipalityDesc == "Langley - City",
         TaxClassCode == "06")
cat("Langley:")
table(Langley$TaxClassTaxRate)

Burnaby <- assessment.thin%>%
  filter(Year == 2019,
         AddressAssessorMunicipalityDesc == "Burnaby",
         TaxClassCode == "06")
cat("Burnaby:")
table(Burnaby$TaxClassTaxRate)

Vancouver <- assessment.thin%>%
  filter(Year == 2019,
         AddressAssessorMunicipalityDesc == "Vancouver",
         TaxClassCode == "01")
cat("Vancouver:")
table(Vancouver$TaxClassTaxRate)

NewWest <- assessment.thin%>%
  filter(Year == 2019,
         AddressAssessorMunicipalityDesc == "New Westminster",
         TaxClassCode == "06")
cat("New Westminster:")
table(NewWest$TaxClassTaxRate)

```
Langley: TaxClassTaxRate also associates with AssetTypeDesc.-> assign 11.8916 due to client request
Delta: tax rate also associate with neighborhood -> Take Mean
Burnaby, Vancovuer and New West: Less that 10 cases are different from others -> Take mode 

```{r}
assessment.rate <- assessment.thin%>%
  mutate(TaxClassTaxRate = case_when(Year %in% c(2019,2020) &
                                       AddressAssessorMunicipalityDesc == "Langley - City" &
                                       TaxClassCode == "06" ~ 11.8916,
                                       Year ==2019 &
                                       AddressAssessorMunicipalityDesc == "Burnaby" &
                                       TaxClassCode == "06" ~ 10.0987,
                                     Year == 2019 &
                                       AddressAssessorMunicipalityDesc == "Vancouver" &
                                       TaxClassCode == "01" ~ 2.56116,
                                     Year == 2019 &
                                       AddressAssessorMunicipalityDesc == "New Westminster" &
                                       TaxClassCode == "06" ~ 14.2158,
                                     TRUE ~ TaxClassTaxRate))
Delta <- assessment.rate%>%
  filter(!is.na(TaxClassTaxRate),
         AddressAssessorMunicipalityDesc == "Delta")%>%
  group_by(AddressAssessorMunicipalityDesc, Year, TaxClassCode)%>%
  summarise(rate = mean(TaxClassTaxRate))

assessment.rate <- merge(assessment.rate, Delta,
                         by = c("AddressAssessorMunicipalityDesc", "Year", "TaxClassCode"), all = TRUE)%>%
  mutate(TaxClassTaxRate = case_when(AddressAssessorMunicipalityDesc == "Delta" ~ rate,
                                     TRUE ~ TaxClassTaxRate))%>%
  select(-rate)

## Check if there is still inconsistancy
(TaxClassTaxRate.text2 <- assessment.rate%>%
    group_by(Year, AddressAssessorMunicipalityDesc,TaxClassCode)%>%
    filter(!is.na(TaxClassTaxRate))%>%
    summarise(test = var(TaxClassTaxRate))%>%
    filter(test !=0))
```
## Impute NA in TaxClassTaxRate
```{r}
assessment.agg <- assessment.rate%>%
  group_by(Year, AddressAssessorMunicipalityDesc,TaxClassCode)%>%
  filter(!is.na(TaxClassTaxRate))%>%
  summarise(rate = unique(TaxClassTaxRate))
assessment.impute <- merge(assessment.rate, assessment.agg,
                           by = c("AddressAssessorMunicipalityDesc", "Year", "TaxClassCode"), all = TRUE)%>%
  mutate(TaxClassTaxRate = rate)%>%
  select(-rate)


(TaxClassTaxRate.text3 <- assessment.impute%>%
    group_by(Year, AddressAssessorMunicipalityDesc,TaxClassCode)%>%
    summarise(test = var(TaxClassTaxRate))%>%
    filter(test !=0))

# List of region that don't have TaxRate
(assessment.miss <- assessment.impute%>%
    filter(is.na(TaxClassTaxRate))%>%
    group_by(Year, AddressAssessorMunicipalityDesc,TaxClassCode)%>%
    summarise(propertyCount = length(unique(PIC))))

# Cleaned data table
assessment.final <- assessment.impute%>%
  filter(!is.na(TaxClassTaxRate))
```
## Aggregate
```{r}
## Aggrehate by Year, AddressAssessorMunicipalityDesc,TaxClassCode
assessment.aggregate <- assessment.final%>%
  group_by(Year, AddressAssessorMunicipalityDesc,TaxClassCode)%>%
  summarise(assessTotal = sum(AssessedValueAmt),
            landTotal = sum(AssessedValueAmt[AssetTypeDesc=="Land"]),
            improvementTotal = sum(AssessedValueAmt[AssetTypeDesc=="Improvement"]),
            propertyCount = length(unique(PIC)),
            rate = unique(TaxClassTaxRate))

## Transform into percentage change
# assessment.pct <- assessment.aggregate%>%
#   group_by(AddressAssessorMunicipalityDesc, TaxClassCode)%>%
#   arrange(Year, .by_group = TRUE) %>%
#   mutate(assessTotal_pct = (assessTotal/lag(assessTotal) - 1) * 100,
#          landTotal_pct = (landTotal/lag(landTotal) - 1) * 100,
#          improvementTotal_pct = (improvementTotal/lag(improvementTotal) - 1) * 100,
#          propertyCount_pct = (propertyCount/lag(propertyCount) - 1) * 100,
#          rate_pct = (rate/lag(rate) - 1) * 100)%>%
#   select(-c(assessTotal,landTotal, improvementTotal, propertyCount, rate))
```
## Export Data

```{r}
write.csv(assessment.final, "../Data/Assessment/assessment_final.csv")
write.csv(assessment.aggregate%>%
            filter(Year != 2020), "../Data/assessment_aggregate.csv")
write.csv(assessment.aggregate%>%
            filter(Year == 2020), "../Data/assessment_2020.csv")
# write.csv(assessment.pct, "../Data/assessment_pct.csv")
```
